R for Clinical Study Reports and Submission

### 2021-10-04 project-folder.Rmd
update running environment.
fix typo

A clearly defined clinical project folder structure can have many benefits to 
clinical development teams in an organization. 
Specifically a well defined project folder can achieve:

- Consistency: every team work on the same folder structure.
- Reproducibility: analysis can be executed and reproduced by different team members months/years later.  
- Automation: automatically check integration of a project.  
- Compliance: reduce compliance issue

We illustrate the R package project folder for A&R project using a demo project called 
[`esubdemo`](https://github.com/elong0527/esubdemo.git). 
You can [clone](https://happygitwithr.com/rstudio-git-github.html#clone-the-new-github-repository-to-your-computer-via-rstudio) the project using RStudio.

For R user, you already benefit from a well defined and consistent folder structure. 
That is the [R package folder structure](https://github.com/rstudio/cheatsheets/raw/master/package-development.pdf). 
Every R package developer is required to follow the same convention to organize their R
functions, before the R package can be disseminated through the Comprehensive R Archive Network
(CRAN). As a user, you can easily install and use those R packages after installing from CRAN. 
There are many good resources to guide developers on R package development.
For example, [Hadley Wickmans' R packages book](https://r-pkgs.org/)

We recommend to use R package folder structure to organize analysis scripts 
for clinical trial development. 
Using the R package folder structure to streamline data analysis work has also been proposed before 
(@marwick2018packaging, @wuanalysis).

@@

For consistency, a well defined folder structure with potential templates
ensures project teams organize A&R work consistently across multiple projects. 
Consistent folder structure also reduce communication cost between study team members 
and enhances the transparency of projects.

In this book, we refer an R package as **project specific R package** 
if the purpose of an R package is to organize analysis scripts for a clinical project.

We refer an R package as **standard R package** if the purpose of an R package is to 
share commonly used R functions to be hosted in an code repository such as CRAN.

Below is a minimal sufficient folders and files for a project specific R package
based on the R package folder structure.

@@

> Here the `adam\` folder contain ADaM datasets. 
> The `tlf` folder contain output TLFs in RTF format.
> We put `adam\` and `tlf\` folder within `vignettes` folder only for illustration purpose. 
> In a real A&R report, you may have different location to save your input and output.

@@

With a formal project, it is also important to provide proper documentations for 
those help functions. We used `Roxygen2` to document functions. 
For example, the header below provide the definition of each variable in `fmt_est`. 
More details can be found in [Chpater 10 of the R package book](https://r-pkgs.org/man.html)

@@

```{r}
# set up snapshot date
snapshot  <- "2021-08-06"

@@

```{r}
source("renv/activate.R")
```

>>>>

A clearly defined clinical project folder structure can have many benefits to 
clinical development teams in an organization. 
Specifically, a well defined project folder can achieve:

- Consistency: everyone work on the same folder structure.
- Reproducibility: analysis can be executed and reproduced by different team members months/years later.  
- Automation: automatically check the integration of a project.  
- Compliance: reduce compliance issue

We illustrate the R package project folder for the A&R project using a demo project called 
[`esubdemo`](https://github.com/elong0527/esubdemo.git). 
You can [clone](https://happygitwithr.com/rstudio-git-github.html#clone-the-new-github-repository-to-your-computer-via-rstudio) the project using RStudio.

For R users, you already benefit from a well-defined and consistent folder structure. 
That is the [R package folder structure](https://github.com/rstudio/cheatsheets/raw/master/package-development.pdf). 
Every R package developer is required to follow the same convention to organize their R
functions before the R package can be disseminated through the Comprehensive R Archive Network
(CRAN). As a user, you can easily install and use those R packages after installing from CRAN. 
There are many good resources to guide developers on R package development.
For example, [Hadley Wickmans' R packages book](https://r-pkgs.org/)

We recommend using the R package folder structure to organize analysis scripts 
for clinical trial development. 
Using the R package folder structure to streamline data analysis work has also been proposed before 
(@marwick2018packaging, @wuanalysis).

@@

For consistency, a well-defined folder structure with potential templates
ensures project teams organize A&R work consistently across multiple projects. 
Consistent folder structure also reduces communication cost between study team members 
and enhances the transparency of projects.

In this book, we refer to an R package as **project specific R package** 
if the purpose of an R package is to organize analysis scripts for a clinical project.

We refer to an R package as a **standard R package** if the purpose of an R package is to 
share commonly used R functions to be hosted in a code repository such as CRAN.

Below is minimal sufficient folders and files for a project specific R package
based on the R package folder structure.

@@

> Here the `adam\` folder contains ADaM datasets. 
> The `tlf` folder contains output TLFs in RTF format.
> We put `adam\` and `tlf\` folder within the `vignettes` folder only for illustration purposes. 
> In a real A&R report, you may have a different location to save your input and output.

@@

With a formal project, it is also important to provide proper documentation for those help functions. 
We used Roxygen2 to document functions. 
For example, the header below defines each variable in fmt_est. 
More details can be found in [Chapter 10 of the R package book](https://r-pkgs.org/man.html)

@@

```{r, eval = FALSE}
# set up snapshot date
snapshot  <- "2021-08-06"

@@

```{r, eval = FALSE}
source("renv/activate.R")
```
